---
title: "MSCI718-Project"
author: "Rushi Raval"
date: "2024-03-24"
output: html_document
---

`{install.packages("brms")} install.packages("dplyr")`

```{r cars}
install.packages("corrplot")

```

## Libraries that will be used further

```{r cars}
library(brms) 
library(corrplot)
library(dplyr)
library(ggplot2)
library(reshape2)
library(stringr)
library(nnet)
library(broom)
library(bayesplot)
library(nnet)
library(dplyr)
library(MASS)
library(ggplot2)
library(lattice)
library(Hmisc)
library(ggcorrplot)
library(gridExtra)
library(memisc)
library(ROCR)
library(rpart)
library(rpart.plot)
library(caret)
library(rms)
library(randomForest)
library(mediation)
```

## Code

Load the csv file:

```{r cars}

file_path <- "Collision.csv"
data <- read.csv(file_path)



print(data)


data <- data[, !(colnames(data) %in% c("X", "Y", "ID", "Geo_ID", "Accident_Date", 
                     "Location", "Num_of_Injuries", "Num_of_Minimal_Injuries", 
                     "Num_of_Minor_Injuries", "Num_of_Major_Injuries", 
                     "Num_of_Fatal_Injuries", "X_Coordinate", "Y_Coordinate", 
                     "Lat", "Long", "ObjectId"))]


```

## Plotting injuries

Code:

```{r cars}

data <- data[data$Max_Injury != "00 - None" & data$Max_Injury != "None" & data$Max_Injury != "", ]


# Create the pie chart for types of injuries
ggplot(data, aes(x = "", fill = Max_Injury)) +
  geom_bar(width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Types of Injuries", 
       fill = "Type of Injury") +
  theme_void()  # Remove unnecessary elements


```

## Plotting accidents with respect to each year

Code:

```{r cars}


ggplot(data, aes(x = factor(Accident_Year))) +
  geom_bar() +
  labs(title = "Distribution of Accidents by Year", 
       x = "Accident_Year",
       y = "Count") +
  scale_x_discrete(labels = unique(data$Accident_Year)) +  # Set x-axis labels to unique years
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better visibility
        plot.title = element_text(hjust = 0.5))


```

## Plotting Accidents with respect to weather conditions

Code:

```{r cars}

data <- data[data$Environment_Condition != "00 - Unknown" ,]


# Create the pie chart for types of injuries
ggplot(data, aes(x = "", fill = Environment_Condition)) +
  geom_bar(width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Types of Environment_Conditions", 
       fill = "Type of Environment_Condition") +
  theme_void()  # Remove unnecessary elements
```

## Plotting accidents with respect to time

Code:

```{r cars}

# Convert the "Time_Range" column to a factor with specified levels

ggplot(data, aes(x = Time_Range)) +
  geom_bar() +  # Use geom_bar() for a bar plot
  labs(x = "Time Range", y = "Count") +  # Add axis labels
  ggtitle("Distribution of Time Ranges") +
  theme(
    text = element_text(size = 20)  # Set the font size to 16
  )# Add plot title


```

## Filter Data and do label encoding

Code:

```{r cars}

categorical_cols <- names(data)[sapply(data, is.character)]

# Convert categorical columns to factors, then to numeric, and subtract 1 to start from 0
data[categorical_cols] <- lapply(data[categorical_cols], function(col) as.numeric(as.factor(col)) - 1)


```
## Correlation matrix

Code:

```{r cars}

head(data)

filled_data <- data
filled_data[, sapply(data, is.numeric)] <- lapply(data[, sapply(data, is.numeric)], function(x) ifelse(is.na(x), mean(x, na.rm = TRUE), x))

# Compute correlation matrix
correlation_matrix <- cor(filled_data[, sapply(filled_data, is.numeric)])

# Plot the heatmap
corrplot(correlation_matrix, method = "color", type = "upper", tl.cex = 0.6, cl.cex = 0.6, number.cex = 0.7, tl.col = "black")

par(cex.main = 1.7, cex.lab = 1.4, cex.axis = 1.4)
corrplot(correlation_matrix, method = "color", type = "upper", tl.cex = 1, cl.cex = 1, number.cex = 1, tl.col = "black")

print(correlation_matrix)

```
## Linear regression and identify useful features

Code:

```{r cars}

lm_model <- lm(Max_Injury ~ Classification_Of_Accident + Time_Range + Location_Type + Initial_Impact_Type + Road_Surface_Condition + Light + Traffic_Control + Environment_Condition, data = data)

step(lm_model)
variable_selection <- step(lm_model)
summary(variable_selection)

```

## Multiple Linear regression

Code:

```{r cars}

lm_model1 <- lm(Max_Injury ~ Road_Surface_Condition + Light + Traffic_Control + Environment_Condition, data = data)

lm_model2 <- lm(Max_Injury ~ Initial_Impact_Type + Road_Surface_Condition + Light + Traffic_Control + Environment_Condition, data = data)

lm_model3 <- lm(Max_Injury ~ Classification_Of_Accident + Time_Range + Initial_Impact_Type + Road_Surface_Condition + Light + Traffic_Control + Environment_Condition, data = data)

lm_model4 <- lm(Max_Injury ~ Location_Type + Time_Range + Classification_Of_Accident + Initial_Impact_Type + Road_Surface_Condition + Environment_Condition + Light + Traffic_Control + Num_of_Vehicle + Num_Of_Pedestrians
, data = data)

mtable(lm_model1, lm_model2, lm_model3, lm_model4,  sdigits = 2)

set.seed(123)  # for reproducibility
train_indices <- sample(nrow(data), 0.7 * nrow(data))  # 70% train, 30% test
train_data <- data[train_indices, ]
test_data <- data[-train_indices, ]

predicted_values <- predict(lm_model4, newdata = test_data)


actual_values <- test_data$Max_Injury

# Mean Absolute Error (MAE)
mae <- mean(abs(predicted_values - actual_values))

# Mean Squared Error (MSE)
mse <- mean((predicted_values - actual_values)^2)

# Root Mean Squared Error (RMSE)
rmse <- sqrt(mse)

cat(mse)






```

## Finding mediator

Code:

```{r cars}

mediator_variable <- data$Environment_Condition

# Define the mediator model
mediator_model <- lm(Road_Surface_Condition ~ Max_Injury + Location_Type + Time_Range + Classification_Of_Accident + Environment_Condition + Road_Surface_Condition + Light + Traffic_Control + Num_of_Vehicle + Num_Of_Pedestrians, data = data)

# Define the outcome model
outcome_model <- lm(Max_Injury ~ Location_Type + Time_Range + Classification_Of_Accident + Initial_Impact_Type + Road_Surface_Condition + Environment_Condition + Light + Traffic_Control + Num_of_Vehicle + Num_Of_Pedestrians + Environment_Condition, data = data)

# Perform mediation analysis
mediation_results <- mediate(mediator_model, outcome_model, treat = "Environment_Condition", mediator="Road_Surface_Condition")

# View the results
summary(mediation_results)

```

## Finding moderator

Code:

```{r cars}

# Assuming you have your lm_model4 already defined
lm_model4 <- lm(Max_Injury ~ Location_Type + Time_Range + Classification_Of_Accident + Initial_Impact_Type + Road_Surface_Condition + Environment_Condition + Light + Traffic_Control + Num_of_Vehicle + Num_Of_Pedestrians, data = data)

# Specify candidate moderator variables (replace "Moderator_Variable" with actual variable names)
candidate_mod_vars <- c("Traffic_Control") 

# Loop through each candidate moderator variable
# Loop through each candidate moderator variable
for (mod_var in candidate_mod_vars) {
  # Create interaction plot
  ggplot(data, aes(x = data[[mod_var]], y = Max_Injury, color = Traffic_Control)) +
    geom_line(aes(group = interaction(Location_Type, data[[mod_var]]))) +
    geom_point() +
    labs(x = mod_var, y = "Traffic_Control", title = paste("Interaction Plot for", mod_var)) +
    theme_minimal()
}


```

## Finding confoudner

Code:

```{r cars}

# Calculate correlation matrix
correlation_matrix <- cor(data[c("Max_Injury", "Location_Type", "Time_Range", "Classification_Of_Accident", "Initial_Impact_Type", "Road_Surface_Condition", "Environment_Condition", "Light", "Traffic_Control", "Num_of_Vehicle", "Num_Of_Pedestrians")])

# Get correlation coefficients with the outcome variable
cor_with_max_injury <- correlation_matrix["Max_Injury", ]

# Identify variables with high correlation coefficients
confounding_variables <- names(which(abs(cor_with_max_injury) > 0.1 & names(cor_with_max_injury) != "Max_Injury"))

# Print potential confounding variables
print(confounding_variables)

```

## Ordinal Logistic Regression

Code:

```{r cars}

data$Max_Injury <- factor(data$Max_Injury, ordered = T)

brm_fit_ord <- brm(Max_Injury ~ Classification_Of_Accident+ Time_Range + Initial_Impact_Type + Road_Surface_Condition + Light + Traffic_Control + Environment_Condition, data = data, family = cumulative(), chains = 1, core = 2, iter = 2000)

summary(brm_fit_ord)

predict(brm_fit_ord)

bayes_R2 (brm_fit_ord)


set.seed(123)  # For reproducibility
train_indices <- sample(nrow(data), 0.8 * nrow(data))  # 70% for training
train_data <- data[train_indices, ]
test_data <- data[-train_indices, ]

predicted_probs <- predict(brm_fit_ord, newdata = test_data, type = "response")

# Convert predicted probabilities to predicted ordinal values
predicted_values <- apply(predicted_probs, 1, function(x) which.max(x) - 1)

# Calculate accuracy
accuracy <- mean(predicted_values == test_data$Max_Injury)

# Print the accuracy
cat("Accuracy:", accuracy, "\n")

```

## Multinomial-Logistic Regression

Code:

```{r cars}

model_glm <- multinom(Max_Injury ~ Location_Type + Classification_Of_Accident + 
    Initial_Impact_Type + Road_Surface_Condition + Environment_Condition + 
    Light + Traffic_Control + Num_of_Vehicle, 
                      data = data)

summary(model_glm)

```

## Multinomial-logistic regression using brms

Code:

```{r cars}


# Convert "Max_Injury" to a factor
class(data$Max_Injury)


# Now try fitting the model again
bayesian_model <- brm(
  formula = Max_Injury ~ Num_of_Vehicle + Initial_Impact_Type + 
    Road_Surface_Condition + Light + Traffic_Control + Environment_Condition, 
  data = data, 
  family = categorical(link = "logit"), 
  chains = 1,           # The number of Markov chains
  cores = 2,
  iter = 2000,# The number of iterations per chain
  warmup = 1000,
  algorithm = "sampling"
)

# Print the summary of the model
summary(bayesian_model)


```

## Decision Tree

Code:

```{r cars}


# Convert "Max_Injury" to a factor
set.seed(123) # for reproducibility

data$Light <- factor(data$Light, levels = unique(data$Light))
data$Road_Surface_Condition <- factor(data$Road_Surface_Condition, levels = unique(data$Road_Surface_Condition))
data$Traffic_Control  <- factor(data$Traffic_Control , levels = unique(data$Traffic_Control ))


train_indices <- createDataPartition(data$Max_Injury, p = 0.7, list = FALSE)
train_data <- data[train_indices, ]
test_data <- data[-train_indices, ]

# Create a decision tree model using the training data
decision_tree <- rpart(Max_Injury ~ Num_of_Vehicle + Initial_Impact_Type + 
    Road_Surface_Condition + Light + Traffic_Control + Environment_Condition,
                       data = train_data,
                       method = "class")

data <- data[data$Light != "00 - Unknown" ,]

# Make predictions on the test data
predictions <- predict(decision_tree, newdata = test_data, type = "class")

# Calculate accuracy
accuracy <- mean(predictions == test_data$Max_Injury)
cat("Accuracy:", accuracy, "\n")

```
